discussion

### 2018-04-05 man/ms_3_discussion.Rmd
update discussion

Although 2005 and 2012 were two severe droughts recorded for the south of the Iberian Peninsula [@GarciaHerrera2007; @VicenteSerrano2014] (Appendix S3), we found a positive trend for greenness of the forests of *Q. pyrenaica* located in their rear edge, as previously was reported using remote sensing vegetation indices [@PerezLuque2015onto; @Alcaraz2016obsnev_ndvi]. For tree-growth, a similar pattern was observed in the last decades, particularly for the southern high-elevation site (CA-High, Figure 4). These results differ from those of @GeaIzquierdo2014 that found a decrease or stabilization in BAI since 1970. Notwithstanding the vulnerability to warming suggested by @GeaIzquierdo2014 for *Q. pyrenaica* forests located at their rear-edge, our results for the same locations, point out a recovery in tree-growth in the last years (Figure 4).

@@

@@

¿Casos parecidos?

@@

trabajos rear edge: 
* sanchez salguero Is drought the main decline factor at the rear edge of Europe? The case of southern Iberian pine plantations 
* sanchez salguero Assessing forest vulnerability to climate warming using a process-based model of tree growth: bad prospects for rear-edges
* dorado-liñán: Climate threats on growth of rear-edge European beech peripheral populations in Spain.
* herrero rigling: Varying climate sensitivity at the dry distribution edge of Pinus sylvestris and P. nigra
* dorado-liñan: Large-scale atmospheric circulation enhances the Mediterranean East-West tree growth contrast at rear-edge deciduous forests. 
buscar otros

Otros estudios que analizan la resiliencia de especies arbóreas frente a los eventos de sequía indican:

@@

Evaluar el crecimiento en el rear edge de F. sylvatica

https://link.springer.com/article/10.1007/s10342-016-0982-7?wt_mc=Internal.Event.1.SEM.ArticleAuthorOnlineFirst

>>>>

Although 2005 and 2012 were two severe droughts recorded for the south of the Iberian Peninsula [@GarciaHerrera2007; @VicenteSerrano2014] (Appendix S3), we found a positive trend for vegetation greenness of the forests of *Q. pyrenaica* located in their rear edge. Our result agrees with those obtained by previous works using other remote sensing vegetation indices [@PerezLuque2015onto; @Alcaraz2016obsnev_ndvi], which suggests an increase in primary productivity during the last years for rear-edge populations of this species.

For tree-growth, a similar positive pattern was observed in the last decades, particularly for the southern high-elevation site (CA-High, Figure 4). This result differs from those previous reported for *Q. pyrenaica* along their distribution range [@GeaIzquierdo2013; @GeaIzquierdo2014]. [@GeaIzquierdo2014] found a general decline in the growth of this species since the 1970s, particularly sharp for rear-edge populations. This decline trend in growth have also been oberved for other tree species located in their rear-edges [*e.g.* @SanchezSalguero2012; @SanchezSalguero2017; @Dorado2017]. Furthermore, growth projections have forecasted a decrease in productivity for *Q. pyrenaica* that would increase vulnerability of this species to climate warming at the dry edges locations [@GeaIzquierdo2013]. However, our results for similar locations have revealed a recovery in tree-growth for the last years (Figure 4).

Las zonas rear edge son zonas que se supone son mas vulnerables al cambio climático por sus estrechos márgenes. De hecho como han visto algunos trabajos, las especies situadas en su rear edge son mas vulnerables. No obstante tambiéh habría que considerar que algunos trabajos apuntan a que la vulnerabilidad dentro del rear edge es mas sitio dependiente (dorado). Además habría que introducir aqui lo que M. Villalta habla de la consideración de rear edge de.

Cuando miramos a la resiliencia de estos bosques podemos ver como aunque en los últimos años ...

Este aumento de BAI, y de greenness no se refleja tanto en la resiliencia observada en los ultimos años. No obstante, n

@@

¿porque? clima + temperatura ...??

@@

```{r echo=FALSE, results='asis', eval=show_text}
cat("
¿Casos parecidos?

@@

trabajos rear edge: 
 * sanchez salguero Is drought the main decline factor at the rear edge of Europe? The case of southern Iberian pine plantations 
 * sanchez salguero Assessing forest vulnerability to climate warming using a process-based model of tree growth: bad prospects for rear-edges
 * dorado-liñán: Climate threats on growth of rear-edge European beech peripheral populations in Spain.
 * herrero rigling: Varying climate sensitivity at the dry distribution edge of Pinus sylvestris and P. nigra
 * dorado-liñan: Large-scale atmospheric circulation enhances the Mediterranean East-West tree growth contrast at rear-edge deciduous forests. 
buscar otros 
")
```

@@

Evaluar el crecimiento en el rear edge de F. sylvatica 
https://link.springer.com/article/10.1007/s10342-016-0982-7?wt_mc=Internal.Event.1.SEM.ArticleAuthorOnlineFirst
